**OptiFit: a fast method for fitting amplicon sequences to existing OTUs**

### 2021-09-27 paper/manuscript.Rmd
final draft
abstract,importance,refs

@@

While comparing AUROC values between models is a useful way to assess the overall model performance, they summarize the performance across all thresholds and can be misleading since models with the same AUROC can have different ROC curve shapes [@lobo2008]. Depending on the intended implementation of the model, one may want to optimize the true-positive rate (sensitivity) over the false-positive rate (1-specificity) or vice versa. In this case, the optimal model will detect as many true positives (people with SRNs) as possible. To further compare the model performance across taxonomic levels we compared the sensitivity of the models at a specificity of 90%. The highest sensitivity values were observed for family, genus, and OTU level data (Figure 1B), consistent with the AUROC results. Phylum, class, order, and ASV sensitivity values were all significantly lower than family, genus, and OTU sensitivity values (Figure 1B). This analysis further supports the observation that finer resolution does not improve SRN detection.

@@

These results demonstrate that consideration of the appropriate taxonomic resolution for utilizing the microbiome as a predictive tool is warranted. In general, we found that finer taxonomic resolution (e.g. OTU and ASV) did not add additional sensitivity to predicting SRNs based on microbiome composition. Family, genus, and OTU level data all performed equally. At the ASV level the fine resolution actually impedes model performance due to the sparsity of shared taxa and led to decreased model performance. The tendency for ASV level annotation to split single bacterial genomes into multiple taxa [@schloss] could also be a contributing factor to the sparsity of shared taxa. Overall, either family, genus, or OTU level taxonomy appear to perform equally for predicting subjects with SRNs based on the composition of the gut microbiome. A potential benefit of utilizing genus or family level data could be that it may allow for merging data generated from different 16S rRNA gene regions or sequencing platforms.

@@

All code is available at: **TODO: link to code**

>>>>

alternative titles:  
High resolution taxonomy impedes microbiome-based classification of colorectal cancer  
High resolution ASVs impede microbiome-based classification of colorectal cancer  
Mid-level taxonomic resolution is optimal for microbiome based classification of colorectal cancer  
Too much of a good thing: high resolution ASVs do not improve microbiome-based classification of colorectal cancer  
More is not always better: high resolution ASVs do not improve microbiome-based classification of colorectal cancer

@@

While comparing AUROC values between models is a useful way to assess the overall model performance, they summarize the performance across all thresholds and can be misleading since models with the same AUROC can have different ROC curve shapes [@lobo2008]. Depending on the intended implementation of the model, one may want to optimize the sensitivity over the specificity or vice versa. In this case, the optimal model will detect as many true positives (people with SRNs) as possible. To further compare the model performance across taxonomic levels we compared the sensitivity of the models at a specificity of 90%. The highest sensitivity values were observed for family, genus, and OTU level data (Figure 1B), consistent with the AUROC results. Phylum, class, order, and ASV sensitivity values were all significantly lower than family, genus, and OTU sensitivity values (Figure 1B). This analysis further supports the observation that finer resolution does not improve SRN detection.

@@

These results demonstrate that consideration of the appropriate taxonomic resolution for utilizing the microbiome as a predictive tool is warranted. In general, we found that finer taxonomic resolution (e.g. OTU and ASV) did not add additional sensitivity to predicting SRNs based on microbiome composition. Family, genus, and OTU level data all performed equally. At the ASV level the fine resolution actually impedes model performance due to the sparsity of shared taxa and led to decreased model performance. The tendency for ASV level annotation to split single bacterial genomes into multiple taxa [@schloss] could also be a contributing factor to the sparsity of shared taxa. Additionally, these results indicate that there are not specific individual bacterial strains that are useful to resolve SRNs, rather sets of closely related bacterial taxa. Overall, either family, genus, or OTU level taxonomy appear to perform equally for predicting subjects with SRNs based on the composition of the gut microbiome. A potential benefit of utilizing genus or family level data could be that it may allow for merging data generated from different 16S rRNA gene regions or sequencing platforms.

@@

All code is available at: https://github.com/SchlossLab/Armour_Resolution_XXXX_2021

### 2021-09-27 paper/manuscript.tex
final draft
abstract,importance,refs

@@

While comparing AUROC values between models is a useful way to assess
the overall model performance, they summarize the performance across all
thresholds and can be misleading since models with the same AUROC can
have different ROC curve shapes (\protect\hyperlink{ref-lobo2008}{9}).
Depending on the intended implementation of the model, one may want to
optimize the true-positive rate (sensitivity) over the false-positive
rate (1-specificity) or vice versa. In this case, the optimal model will
detect as many true positives (people with SRNs) as possible. To further
compare the model performance across taxonomic levels we compared the
sensitivity of the models at a specificity of 90\%. The highest
sensitivity values were observed for family, genus, and OTU level data
(Figure 1B), consistent with the AUROC results. Phylum, class, order,
and ASV sensitivity values were all significantly lower than family,
genus, and OTU sensitivity values (Figure 1B). This analysis further
supports the observation that finer resolution does not improve SRN
detection.

@@

These results demonstrate that consideration of the appropriate
taxonomic resolution for utilizing the microbiome as a predictive tool
is warranted. In general, we found that finer taxonomic resolution
(e.g.~OTU and ASV) did not add additional sensitivity to predicting SRNs
based on microbiome composition. Family, genus, and OTU level data all
performed equally. At the ASV level the fine resolution actually impedes
model performance due to the sparsity of shared taxa and led to
decreased model performance. The tendency for ASV level annotation to
split single bacterial genomes into multiple taxa
(\protect\hyperlink{ref-schloss}{12}) could also be a contributing
factor to the sparsity of shared taxa. Overall, either family, genus, or
OTU level taxonomy appear to perform equally for predicting subjects
with SRNs based on the composition of the gut microbiome. A potential
benefit of utilizing genus or family level data could be that it may
allow for merging data generated from different 16S rRNA gene regions or
sequencing platforms.

@@

All code is available at: \textbf{TODO: link to code}

>>>>

alternative titles:\\
High resolution taxonomy impedes microbiome-based classification of
colorectal cancer\\
High resolution ASVs impede microbiome-based classification of
colorectal cancer\\
Mid-level taxonomic resolution is optimal for microbiome based
classification of colorectal cancer\\
Too much of a good thing: high resolution ASVs do not improve
microbiome-based classification of colorectal cancer\\
More is not always better: high resolution ASVs do not improve
microbiome-based classification of colorectal cancer

@@

While comparing AUROC values between models is a useful way to assess
the overall model performance, they summarize the performance across all
thresholds and can be misleading since models with the same AUROC can
have different ROC curve shapes (\protect\hyperlink{ref-lobo2008}{9}).
Depending on the intended implementation of the model, one may want to
optimize the sensitivity over the specificity or vice versa. In this
case, the optimal model will detect as many true positives (people with
SRNs) as possible. To further compare the model performance across
taxonomic levels we compared the sensitivity of the models at a
specificity of 90\%. The highest sensitivity values were observed for
family, genus, and OTU level data (Figure 1B), consistent with the AUROC
results. Phylum, class, order, and ASV sensitivity values were all
significantly lower than family, genus, and OTU sensitivity values
(Figure 1B). This analysis further supports the observation that finer
resolution does not improve SRN detection.

@@

These results demonstrate that consideration of the appropriate
taxonomic resolution for utilizing the microbiome as a predictive tool
is warranted. In general, we found that finer taxonomic resolution
(e.g.~OTU and ASV) did not add additional sensitivity to predicting SRNs
based on microbiome composition. Family, genus, and OTU level data all
performed equally. At the ASV level the fine resolution actually impedes
model performance due to the sparsity of shared taxa and led to
decreased model performance. The tendency for ASV level annotation to
split single bacterial genomes into multiple taxa
(\protect\hyperlink{ref-schloss}{12}) could also be a contributing
factor to the sparsity of shared taxa. Additionally, these results
indicate that there are not specific individual bacterial strains that
are useful to resolve SRNs, rather sets of closely related bacterial
taxa. Overall, either family, genus, or OTU level taxonomy appear to
perform equally for predicting subjects with SRNs based on the
composition of the gut microbiome. A potential benefit of utilizing
genus or family level data could be that it may allow for merging data
generated from different 16S rRNA gene regions or sequencing platforms.

@@

All code is available at:
\url{https://github.com/SchlossLab/Armour_Resolution_XXXX_2021}
